$K$-corrections and filter transformations \\ from the restframe ultraviolet to the near infrared

### 2005-09-07 docs/paper/kcorrect.tex
update docs and paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@504 75dff457-8751-0410-ab1a-cd701654ce6b
update paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@520 75dff457-8751-0410-ab1a-cd701654ce6b
paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@523 75dff457-8751-0410-ab1a-cd701654ce6b

\author{
Michael R. Blanton\altaffilmark{\ref{NYU}} and 
Samuel Roweis
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%J.~Brinkmann\altaffilmark{\ref{APO}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Mamoru Doi\altaffilmark{\ref{Tokyo}},
%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%James E. Gunn\altaffilmark{\ref{Princeton}},
%David W. Hogg\altaffilmark{\ref{NYU}}, and
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

From the spectroscopic observations we know that galaxy spectra reside
in a low-dimensional locus. Principal Component Analysis (PCA) of
galaxy spectra, introduced by \citet{connolly94a} and applied many
times since then ({\bf more refs}; \citealt{yip05a}), demonstrates
that most of the variance in the distribution of galaxies in spectral
space can be explained using a few templates. This means that, in the
linear space of all possible spectra, galaxies exist in only a small
locus. Therefore, even with very heterogeneous data, we should be able
to determine the properties of this locus.

@@

\subsection{The models}

@@

Note that there are 20 different broadband photometric filters listed
above ($B$ in DEEP2 is different than the $B$ used by GOODS). The {\tt
kcorrect} product described in Section \ref{code} contains a
tabulation of the response functions for all of these filters.

@@

Figure \ref{lrg_colors} shows the LRG colors as a function of
redshift, with the best fit template color overplotted as the smooth,
thick line. In the code described in Section \ref{code}, we used the
routine {\tt sdss\_kcorrect} (with the optional flag {\tt /lrg}) to
perform these fits. The best fit is a good fit to the data.

@@

These templates explain the photometric data rather well. For example,
consider Figure \ref{fullfits}, showing the color residuals of the
observations with respect to the best fits, when fitting to galaxies
with GALEX, SDSS, and 2MASS data. In this context we define the color
residuals as (for example):
\begin{equation}
\Delta [u-g] = [u_{\mathrm{obs}}- g_{\mathrm{obs}}] -
[u_{\mathrm{model}}- g_{\mathrm{model}}] .
\end{equation}
In the code described in Section \ref{code}, we used the routine {\tt
gst\_kcorrect} to perform these fits.  In Figure \ref{fullfits} the
thick dashed lines show the estimated 1$\sigma$ uncertainties in the
colors from the photometric catalogs. Relative to the uncertainties,
there are no significant biases or redshift trends in these fits. Note
that the galaxies used in this test are distinct from the training set
on which we fit the five templates.

@@

A particularly common special case is when $R=Q$:
\begin{equation}
\label{eq:specialL}
K_{QR} = -2.5\,\log_{10}\left[\frac{1}{[1+z]}\,
  \frac{\displaystyle
  \int\mathrm{d}\lambdaobs\,\lambdaobs\,L_{\lambda}\!\left(\frac{\lambdaobs}{1+z}\right)\,R(\lambdaobs)\,}
       {\displaystyle
    \int\mathrm{d}\lambdaemit\,\lambdaemit\,
    L_{\lambda}(\lambdaemit)\,R(\lambdaemit)}
\right] \;\;\;.
\end{equation}

@@

\citet{kauffmann03a} have also calculated stellar masses for SDSS
galaxies. Roughly, they have calculated the $z$-band stellar
mass-to-light ratios of models which best fit the H$\delta$ absorption
and D4000 measurements of the spectra, and then applied those
mass-to-light ratios to the observed $z$-band luminosities of the
imaging. Figure \ref{mass_to_garching} shows a galaxy-by-galaxy
comparison of their stellar masses $M_{\ast,s}$ relative to ours
$M_\ast$, as a function of $M_\ast$ and of intrinsic $g-r$ color. The
two sets of masses are very similar to each other, with a scatter of
only 0.1 dex and trends of less than 0.2 dex with stellar mass. As we
did, \citet{kauffmann03a} used the library of \citet{bruzual03a} with
the \citet{chabrier03} initial mass function, which makes some of this
agreement unsurprising.  To be explicit, the masses in Figure
\ref{mass_to_garching} correspond to the current mass in stars
remaining in the galaxy. This mass can differ from the total
star-formation rate integrated over time, due to the fact that stars
can die --- and in practice for our fits the total integrated
star-formation rate is usually about twice as high as the total
current mass in stars.

Similarly, Figure \ref{mtol} shows the mass-to-light ratios of SDSS
galaxies in the $V$ band as a function of color in $B-V$. Here, we
have calculated the $B$ and $V$ band fluxes based on the template fits
to each galaxy, as explained in Section \ref{sdss2bessell} below. We
give the mass-to-light ratio in solar units. The solid line is the
relationship that \citet{bell01a} give based on their fits to spiral
galaxies. For blue galaxies ($B-V < 0.8$) this line is indeed a good
fit to our results. For the typical red galaxy, the \citet{bell01a}
relation predicts a larger stellar mass than we do.

Finally, we can also calculate measures of the recent star-formation
rate from these fits. Since our fits are to broad-band photometry, and
not based on emission line measurements, we cannot expect to have
detailed information about the very recent star-formation
rate. However, we can try to measure what fraction of the total
star-formation has occurred in, say, the past 1 Gyr:
\begin{equation}
b_{G} = \frac{\int_0^{1\mathrm{~Gyr}} dt\, \mathrm{SFR}(t)}
{\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
\end{equation}
Note that here $b_G$ is defined in terms of the total integrated
star-formation history (unlike the stellar masses referred to above)
as is standard for the $b$ parameter (\citealt{kennicutt98a}). Figure
\ref{umr_bg} shows the relationship between the rest-frame $u-r$ color
for SDSS galaxies and $b_G$. For galaxies with $u-r < 2.5$ (redder
than which galaxies are mostly reddened, edge-on disks) these two
parameters follow a simple relationship:
\begin{equation}
\log_{10} b_g = -0.55 (u-r) - 0.25.
\end{equation}

@@

However, often one wants to just make quick and dirty comparisons,
without going through the trouble of running any software. For these
purposes, we provide Table \ref{lineareqs}, which provides the linear
relationships among the same bandpasses as those listed in Table
\ref{solarmagnitudes}. Note that Table \ref{lineareqs} refers to AB
magnitudes {\it throughout}; use Table \ref{solarmagnitudes} to get
the relationships to Vega magnitudes.

[ sdss2bessell ]

[ conversions ]

\section{Public access to the code}
\label{code}

\section{Summary}
\label{conclusions}

[ summarize ]

[ this does not describe what constraints photometry puts on the SFH ]

@@

@@

@@

\section{$K$-corrections to shifted bandpasses}
\label{bandshift}

>>>>

\author{
Michael R. Blanton\altaffilmark{\ref{NYU}} and 
Sam Roweis
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%J.~Brinkmann\altaffilmark{\ref{APO}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Mamoru Doi\altaffilmark{\ref{Tokyo}},
%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%James E. Gunn\altaffilmark{\ref{Princeton}},
%David W. Hogg\altaffilmark{\ref{NYU}}, and
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

From the spectroscopic observations we know that galaxy spectra reside
in a low-dimensional locus. Principal Component Analysis (PCA) of
galaxy spectra, introduced by \citet{connolly95b} and applied many
times since then, demonstrates that most of the variance in the
distribution of galaxies in spectral space can be explained using a
few templates. This means that, in the linear space of all possible
spectra, galaxies exist in only a small locus. Therefore, even with
very heterogeneous data, we should be able to determine the properties
of this locus.

@@

\subsection{The models}
\label{models}

@@

Note that there are 20 different broadband photometric filters listed
above ($B$ in DEEP2 is different than the $B$ used by GOODS). The {\tt
kcorrect} product described in Section \ref{summary} contains a
tabulation of the response functions for all of these filters.

@@

Figure \ref{lrg_colors} shows the LRG colors as a function of
redshift, with the best fit template color overplotted as the smooth,
thick line. In the code described in Section \ref{summary}, we used the
routine {\tt sdss\_kcorrect} (with the optional flag {\tt /lrg}) to
perform these fits. The best fit is a good fit to the data.

@@

These templates explain the photometric data rather well. For example,
consider Figure \ref{fullfits}, showing the color residuals of the
observations with respect to the best fits, when fitting to galaxies
with GALEX, SDSS, and 2MASS data. In this context we define the color
residuals as (for example):
\begin{equation}
\Delta [u-g] = [u_{\mathrm{obs}}- g_{\mathrm{obs}}] -
[u_{\mathrm{model}}- g_{\mathrm{model}}] .
\end{equation}
In the code described in Section \ref{summary}, we used the routine {\tt
gst\_kcorrect} to perform these fits.  In Figure \ref{fullfits} the
thick dashed lines show the estimated 1$\sigma$ uncertainties in the
colors from the photometric catalogs. Relative to the uncertainties,
there are no significant biases or redshift trends in these fits. Note
that the galaxies used in this test are distinct from the training set
on which we fit the five templates.

@@

A particularly common special case is when $R=Q$:
\begin{equation}
\label{eq:specialL}
K_R(z) = -2.5\,\log_{10}\left[\frac{1}{[1+z]}\,
  \frac{\displaystyle
  \int\mathrm{d}\lambdaobs\,\lambdaobs\,L_{\lambda}\!\left(\frac{\lambdaobs}{1+z}\right)\,R(\lambdaobs)\,}
       {\displaystyle
    \int\mathrm{d}\lambdaemit\,\lambdaemit\,
    L_{\lambda}(\lambdaemit)\,R(\lambdaemit)}
\right] \;\;\;.
\end{equation}

@@

\citet{kauffmann03a} have also calculated stellar masses for SDSS
galaxies. Roughly, they have calculated the $z$-band stellar
mass-to-light ratios of models which best fit the H$\delta$ absorption
and D4000 measurements of the spectra, and then applied those
mass-to-light ratios to the observed $z$-band luminosities of the
imaging. Figure \ref{mass_to_garching} shows a galaxy-by-galaxy
comparison of their stellar masses $M_{\ast,s}$ relative to ours
$M_\ast$, as a function of $M_\ast$ and of intrinsic $g-r$ color. The
two sets of masses are very similar to each other, with a scatter of
only 0.1 dex and trends of less than 0.2 dex with stellar mass. As we
did, \citet{kauffmann03a} used the library of \citet{bruzual03a} with
the \citet{chabrier03a} initial mass function, which makes some of this
agreement unsurprising.  To be explicit, the masses in Figure
\ref{mass_to_garching} correspond to the current mass in stars
remaining in the galaxy. This mass can differ from the total
star-formation rate integrated over time, due to the fact that stars
can die --- and in practice for our fits the total integrated
star-formation rate is usually about twice as high as the total
current mass in stars.

Similarly, Figure \ref{mtol} shows the mass-to-light ratios of SDSS
galaxies in the $V$ band as a function of color in $B-V$. Here, we
have calculated the $B$ and $V$ band fluxes based on the template fits
to each galaxy, as explained in Section \ref{sdss2bessell} below. We
give the mass-to-light ratio in solar units. The solid line is the
relationship that \citet{bell01b} give based on their fits to spiral
galaxies. For blue galaxies ($B-V < 0.8$) this line is indeed a good
fit to our results. For the typical red galaxy, the \citet{bell01b}
relation predicts a larger stellar mass than we do.

Finally, we can also calculate measures of the recent star-formation
rate from these fits. Since our fits are to broad-band photometry, and
not based on emission line measurements, we cannot expect to have
detailed information about the very recent star-formation
rate. However, we can try to measure what fraction of the total
star-formation has occurred in, say, the past 1 Gyr:
\begin{equation}
b_{G} = \frac{\int_0^{1\mathrm{~Gyr}} dt\, \mathrm{SFR}(t)}
{\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
\end{equation}
Note that here $b_G$ is defined in terms of the total integrated
star-formation history (unlike the stellar masses referred to above)
as is standard for the $b$ parameter (\citealt{kennicutt98a}). Figure
\ref{umr_bg} shows the relationship between the rest-frame $u-r$ color
for SDSS galaxies and $b_G$. For galaxies with $u-r < 2.5$ (redder
than which galaxies are mostly reddened, edge-on disks) these two 
parameters follow a simple relationship:
\begin{equation}
\log_{10} b_g = -0.55 (u-r). 
\end{equation}

@@

However, often one wants to just make quick and dirty comparisons,
without going through the trouble of running any software. For these
purposes, we provide Table \ref{lineareqs}, which provides the linear
relationships among the same bandpasses as those listed in Table
\ref{solarmagnitudes}. In fact, these relationships are usually good
to 0.05 mag or better. Note that Table \ref{lineareqs} refers to AB
magnitudes {\it throughout}; use Table \ref{solarmagnitudes} to get
the relationships to Vega magnitudes.

\section{Summary}
\label{summary}

We have here described a method for fitting templates to
nearly-arbitrary sets of spectra and broad-band fluxes. The basic
concept is that we can recover a small set of templates (based on
models), nonnegative linear combinations of which can explain a much
larger set of inhomogeneous observations.  We have applied this method
to SDSS, GALEX, 2MASS, DEEP2, and GOODS data to come up with a set of
templates which are effective for describing all of these data sets.

Nonnegative matrix factorization (NMF) is an effective way to reduce
the dimensionality of any large data set, and has that in common with
PCA.  It is, in some sense, a ``nonnegative PCA.''  However, at least
in the current context the method has several advantages over
PCA. First, it has a natural physical interpretation associated with
that spectral subspace, which is the corresponding subspace of all
possible star-formation histories. Second, it naturally handles data
uncertainties and missing data, which allows it to ignore that
variation which is due purely to statistical errors.  Third, it
handles the complications of observing galaxy spectra photometrically
using broad-band filters.

We note here that the general NMF method does not {\tt depend} on
using a model --- our ``model'' could just have been a set of top hat
functions or Gaussians on a wavelength grid. In some situations, such
as datasets for which there are not well-developed theoretical models,
this approach could be more appropriate.

We have released our results in the form of a templates and a code
base called {\tt kcorrect} which fits those templates to many types of
data (SDSS photometric and spectroscopic data, GALEX data, GOODS data,
DEEP2 data, and 2MASS data). Furthermore, this code returns
$K$-corrections and a physical interpretation of the photometry. All
of the plots in this paper were created using code in the
repository. The code is available from a web page maintained by one of
the authors\footnote{{\tt
http://cosmo.nyu.edu/blanton/kcorrect}}. This web page is kept updated
on improvements in the code and new developments. It consists of a
{\tt C} language library, stand-alone {\tt C} programs, and IDL
language wrappers around the {\tt C} library. Thus, one can use the
basic templates and fitting code using only a machine with a {\tt C}
compiler. However, there is significant functionality which is
programmed in the (unfortunately, proprietary) IDL language. These
routines depend on the {\tt idlutils} library\footnote{{\tt
http://skymaps.info}}.

@@

The authors would like to thank Alison Coil, David W. Hogg, David
Schiminovich,

@@

[sam sam sam sam]

@@

We have fit several different sets of templates that we release
with the code, which we denote:
\begin{enumerate}
\item {\tt default}, the default set of five templates, 
\item {\tt lrg1}, the single template fit to LRGs, and 
\item {\tt goods}, the five template fit to just the GOODS data.
\end{enumerate}
The information about the default set of templates is contained in the
file {\tt data/templates/k\_nmf\_derived.default.fits} in the {\tt
kcorrect} project.  This file has 25 Header and Data Units (HDUs),
each listed in Table \ref{derived} and described in more detail in the
paragraphs below.  There are similar files for the {\tt lrg1} and {\tt
goods} template sets.

As described in Section \ref{models}, there are $N_{\mathrm{basis}} =
485$ basis spectra, consisting of 450 different instantaneous burst
stellar populations ($N_{\mathrm{age}} = 25$ ages, $N_{\mathrm{mets}}
= 6$ metallicities, and $N_{\mathrm{dust}}=3$ dust properties) plus 35
different emission line models. The {\tt templates} HDU has the
coefficients of the five templates in this basis space.

For each of the five templates, we have the template spectrum {\tt
spec} in units of ergs s$^{-1}$ cm$^{-2}$ \AA$^{-1}$ per solar mass,
as it would be observed at 10 pc distance. The wavelength grid is in
the {\tt lambda} HDU. As noted above, the models are smoothed at 300
km s$^{-1}$ resolution; the {\tt rawspec} HDUs have the original
(unsmoothed) models from \citet{bruzual03a}. In addition, we give
versions without the emission lines (with the {\tt \_nl} suffix) and
without the dust extinction applied (with the {\tt \_nd} suffix). The
actual emission-line-only spectra for each template are in the {\tt
lspec} HDU. Finally, the multiplicative dust extinction factor for
each template spectrum is given in the {\tt extinction} HDU.

We include the star-formation rate as a function of time in the {\tt
sfr} HDU (the age grid used is in the {\tt ages} HDU). The time
differential used to quantify this rate is in the {\tt dage} HDU; to
get the total number of stars formed at each age, multiply {\tt sfr}
by {\tt dage}. The average metallicity of the stars formed as a
function of time is in the {\tt metallicity} HDU.

The total mass formed in each template (in solar masses) is in the
{\tt mass} HDU --- this is actually just unity for each template. The
total surviving stellar mass for each template is in the {\tt mremain}
HDU. The metallicity in the surviving stars is given in the {\tt mets}
HDU.  The total mass in stars formed in the last 300 Myrs is in the
{\tt m300} HDU, and the total mass formed in the last 1 Gyr is in the
{\tt m1000} HDU.

The properties of the 450 stellar population basis vectors are given
in the four HDU: {\tt basis\_ages}, {\tt basis\_mets}, {\tt
basis\_dusts}, and {\tt basis\_mremain}. Respectively, these give the
ages, metallicities, dust properties, and fraction of original stellar
mass remaining for each instantaneous burst in the grid. The dust 
properties are given in terms of a structure with four elements which
refer to the properties in the multiple scattering model of \citet{witt00a}:
\begin{enumerate}
\item {\tt GEOMETRY}: the geometry of the dust (e.g. ``shell'') 
\item {\tt DUST}: the dust extinction curve (e.g. ``MW'' for Milky
	Way extinction and ``SMC'' for Small Magellanic Cloud type
	extinction) 
\item {\tt STRUCTURE}: structure of the dust distribution (``h'' for
	homogeneous, ``c'' for clumpy)
\item {\tt TAUV}: the amount of dust (the total optical depth in the
	$V$ band)
\end{enumerate}

### 2005-09-07 docs/paper/kcorrect_figures.tex
paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@523 75dff457-8751-0410-ab1a-cd701654ce6b

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{spec_lrg.ps}
\caption{\label{spec_lrg} Best fit LRG spectral template
	(normalization is for a 1 $M_\odot$ galaxy located 10 pc away, or
	equivalently a $10^12$ $M_\odot$ galaxy located 10 Mpc away).}
\end{figure}

@@

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{mtol.ps}
\caption{\label{mtol} Mass-to-light ratios of galaxies in the $V$ band
	(in solar units) as a function of galaxy $B-V$ color. The solid line
	satisfies the relationship $\log_{10}(M/L_V) = 1.40 (g-r) - 0.73$,
	given by \citet{bell01a} for their sample of spiral galaxies. 
	Their estimates and ours agree for $B-V < 0.8$, where spiral
	galaxies dominate the galaxy population.}
\end{figure}

>>>>

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{spec_lrg.ps}
\caption{\label{spec_lrg} Best fit LRG spectral template
	(normalization is for a 1 $M_\odot$ galaxy located 10 pc away, or
	equivalently a $10^{12}$ $M_\odot$ galaxy located 10 Mpc away).}
\end{figure}

@@

\clearpage
\stepcounter{thefigs}
\begin{figure}
\figurenum{\fignum}
\plotone{mtol.ps}
\caption{\label{mtol} Mass-to-light ratios of galaxies in the $V$ band
	(in solar units) as a function of galaxy $B-V$ color. The solid line
	satisfies the relationship $\log_{10}(M/L_V) = 1.40 (g-r) - 0.73$,
	given by \citet{bell01b} for their sample of spiral galaxies. 
	Their estimates and ours agree for $B-V < 0.8$, where spiral
	galaxies dominate the galaxy population.}
\end{figure}

### 2005-09-07 docs/paper/kcorrect_tables.tex
update paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@520 75dff457-8751-0410-ab1a-cd701654ce6b
paper
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@523 75dff457-8751-0410-ab1a-cd701654ce6b

\begin{deluxetable}{cc}
\tablewidth{0pt}
\tablecolumns{2}
\tablecaption{\label{lineareqs} Conversions among various filters}
\tablehead{ Equation & Color dispersion }
\startdata
\input{lineareqs.tex}
\enddata
\tablecomments{Uses AB magnitudes throughout.}
\end{deluxetable}

>>>>

\begin{deluxetable}{lr}
\tablewidth{0pt}
\tablecolumns{2}
\tablecaption{\label{lineareqs} Conversions among various filters}
\tablehead{ Equation & Color dispersion }
\startdata
\input{lineareqs.tex}
\enddata
\tablecomments{Uses AB magnitudes throughout.}
\end{deluxetable}

\begin{deluxetable}{llcl}
\tablewidth{0pt}
\tablecolumns{4}
\tablecaption{\label{derived} HDU of {\tt k\_nmf\_derived.default.fits}}
\tablehead{ Number & Name & Dimensions & Description}
\startdata
0 & {\tt templates} & $N_{\mathrm{basis}} \times N_t$ & Coefficients
in basis space for each template \cr
1 & {\tt spec} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template \cr
2 & {\tt spec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template, lines removed \cr
3 & {\tt spec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template, without dust \cr
4 & {\tt spec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ & Smoothed spectrum of
each template, without lines or dust \cr
5 & {\tt rawspec} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
each template \cr
6 & {\tt rawspec\_nl} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
each template, lines removed \cr
7 & {\tt rawspec\_nd} & $N_{\mathrm{spec}} \times N_t$ & Unsmoothed spectrum of
each template, without dust \cr
8 & {\tt rawspec\_nl\_nd} & $N_{\mathrm{spec}} \times N_t$ &
Unsmoothed spectrum of each template, without lines or dust \cr
9 & {\tt lspec} & $N_{\mathrm{spec}} \times N_t$ & emission line spectrum \cr
10 & {\tt extinction} & $N_{\mathrm{spec}} \times N_t$ & extinction as
a function of wavelength \cr
11 & {\tt lambda} & $N_{\mathrm{spec}}$ & Wavelength grid (\AA) \cr
12 & {\tt sfr} & $N_{\mathrm{age}} \times N_t$ & Star-formation rate
($M_\odot$ per year) as a function of time \cr
13 & {\tt metallicity} & $N_{\mathrm{age}} \times N_t$ & Average
metallicity as a function of time \cr
14 & {\tt ages} & $N_{\mathrm{age}}$ & Age grid (yrs) \cr
15 & {\tt dage} & $N_{\mathrm{age}}$ & Age bin size (yrs);
multiply {\tt sfr} for total star-formation in each bin \cr
16 & {\tt mass} & $N_t$ & total mass formed in each template \cr
17 & {\tt mremain} & $N_t$ & total current stellar mass in each template \cr
18 & {\tt mets} & $N_t$ & average metallicity of current stars in each
template \cr
19 & {\tt m300} & $N_t$ & total stellar mass formed in last 300 Myrs in each template \cr
20 & {\tt m1000} & $N_t$ & total stellar mass formed in last 1 Gyr in each template \cr
21 & {\tt basis\_ages} &
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
ages of each stellar population basis vector  \cr
22 & {\tt basis\_mets} & 
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
metallicity of each stellar population basis vector  \cr
23 & {\tt basis\_dusts} & 
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
dust properties of each stellar population basis vector \cr
24 & {\tt basis\_mremain} & 
$N_{\mathrm{age}} \times N_{\mathrm{mets}} \times N_{\mathrm{dust}}$ &
fraction of original stellar mass surviving for age of each basis vector  \cr
\enddata 
\tablecomments{All quantities are in floating point (four byte)
precision, except {\tt dusts} which is an array of structures, whose
format is described in text.}
\end{deluxetable}
